---
title: "Lab 6"
author: "Lauren Harris"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Rank-based test example

Mann-Whitney U (non-parametric alternative to an unpaired 2-sample t-test)
- H0: equal ranks (or medians)
- HA: ranks are not equal (or medians are not equal)

### Create mock samples

`set.seed()` to create "pseudo-random but reproducible samples"
```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

Take a quick look at the distributions using `hist()`

```{r}
hist(gp_1)
hist(gp_2)
```

Try a t-test so we can compare the results
```{r}
ttest <- t.test(gp_1, gp_2)
```

This p-value means that if our samples are drawn from populations with the same mean (i.e. if the null is true) there is a `r round(ttest$p.value, 2)` probability of finding sample means *at least as different* as mine by random chance.

### Mann Whitney U rank-based test

Use the `wilcox.test()` function to perform a two-sample rank-based test
```{r}
mwu <- wilcox.test(gp_1, gp_2)
```
This p-value means that if our samples are drawn from a population with the same median, there is a `r round(mwu$p.value, 2)` probability of finding samples medians at least as different as ours by random chance

No significant difference in medians (Mann Whitney U; U = `r mwu$statistic`, p-value = `r round(mwu$p.value, 2)`)

## Simple linear regression

Two *continuous quantitative* variables you are trying to understand the relationship between, and you think that a linear relationship makes sense

Use `penguins` data from {palmerpenguins}, exploring flipper length (x) versus body mass (y)

```{r}
# Use View(penguins) to see the data
# Make an exploratory scatter plot of penguin flipper length vs body mass 
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point() +
  geom_smooth(method = "lm")
```

### Find the model estimates (coefficients)

Use the `lm()` funciton for a linear model
```{r}
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)
```

`r penguin_lm$coefficients[1]`

### Use `broom::tidy()` to get a tidy output from lm()
```{r}
penguin_lm_tidy <- tidy(penguin_lm)

penguin_lm_tidy
```

### Use `broom::glance()` to get output of model fit and statistical significance
```{r}
penguin_lm_glance <- glance(penguin_lm)

penguin_lm_glance
```

### Use equatiomatics::extract_eq() to return the model in LaTeX

```{r}
extract_eq(model = penguin_lm, use_coefs = TRUE)
```

### Use `plot()` function to automatically produce diagnostic plots

```{r}
plot(penguin_lm)
```

